---
title: "p8105_hw1_ll3255"
author: "LML"
date: "9/16/2019"
output: html_document
---

```{r, eval= FALSE}
library(tidyverse)
```

#__Problem 1:__ 
Here's a code chunk that samples from a normal distribution:
```{r}
hw1_df = tibble(
  prob1_samp = rnorm(8),
  prob1_positive = prob1_samp > 0,
  prob1_logic = prob1_positive > 0,
  prob1_char = c("one", "two", "three", "four", "five", "six", "seven", "eight"), 
  prob1_factor = factor(prob1_char, levels = c("low", "moderate", "high")),
)
```



Here's a code chunk that is attempting to take the mean of each variable in the data frame above:
```{r}
mean(pull(hw1_df, prob1_samp))
mean(pull(hw1_df, prob1_positive))
mean(pull(hw1_df, prob1_logic))
mean(pull(hw1_df, prob1_char))
mean(pull(hw1_df, prob1_factor))
```
  prob1_samp WORKS
  prob1_positive
  prob1_logic 
  prob1_char 
  prob1_factor



Here's a code chunk is attempting to convert variable types:
## Logical to numeric
```{r}
coerce_log_num = as.numeric(pull(hw1_df, prob1_logic))
```
## Character to numeric
```{r, eval= FALSE}
coerce_char_num = as.numeric(pull(hw1_df, prob1_char))
```
## Factor to numeric
```{r}
coerce_fac_num = as.numeric(pull(hw1_df, prob1_factor))
```
* What happens and why: Logical to numeric became binary 1s and 0s/ Character to numeric became all "NA". Factor to numeric also became all "NA".
* Does this explain what happenned to the mean before: The mean was being calculated from the binary coded variables of 1 and 0. 



Here's a second code chunk converting and manipulating the variables further:
## 1. converting the logical vector to numeric and multiplying the random sample by the result
```{r}
coerce_log_num*pull(hw1_df, prob1_samp)
```
## 2. **FIX THIS LAURA: converting the logical vector to a factor and multiplying the random sample by the result**
```{r}
coerce_log_fac = as.factor(pull(hw1_df, prob1_logic))
  coerce_log_fac*pull(hw1_df, prob1_samp)
```
Result: Not meaningful factors, all are NA

## 3. **FIX THIS LAURA: converting the logical vector to a factor and then to numeric and multiplying the random sample by the result**
```{r}
coerce_log_fac = as.factor(pull(hw1_df, prob1_logic))
  coerce_fac_num = as.numeric(coerce_log_fac)
  coerce_fac_num*pull(hw1_df,  prob1_samp)
```
Result: Get numeric variables


#__Problem 2:__
Here's a code chunk creating a new data frame:
```{r}
prob2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  prob2_logic = x + y > 1,
  log_to_num = as.numeric(prob2_logic),
  log_to_fac = as.factor(prob2_logic)
)
```



Descriptive statistics of dataset "prob2_df":
There are `r nrow(prob2_df)` rows and `r ncol(prob2_df)` columns in the dataset.
The mean of x is `r mean(pull(prob2_df, x))`
The median of the x is `r median(pull(prob2_df, x))`
The standard deviation of the x is `r median(pull(prob2_df, x))`
The proportion of cases f for which x + y >1 is `r sum(pull(prob2_df, prob2_logic= "true")/(nrow(prob2_df))`



Here's a code chunk of a scatterplot for y vs. x with logical color points:
```{r}
ggplot(prob2_df, aes(x = x, y = y)) + geom_point(aes (color = prob2_logic)) + labs(title ="X vs. Y by Logic Function x + y >1")
```



This one has numeric color points:
```{r}
ggplot(prob2_df, aes(x = x, y = y)) + geom_point(aes (color = log_to_num)) + labs(title ="X vs. Y by Numeric Variable")
```
This one has factor color points:
```{r}
ggplot(prob2_df, aes(x = x, y = y)) + geom_point(aes (color = log_to_fac)) + labs(title ="X vs. Y by Factor Variable")
```

Finally export the first scatterplot to the project directory using ggsave.

