---
title: "p8105_hw1_ll3255"
author: "LML"
date: "9/16/2019"
output: html_document
---

```{r}
library(tidyverse)
```

#__Problem 1:__ 
Here's a **code chunk** that samples from a normal distribution:
```{r}
hw1_df = tibble(
  prob1_samp = rnorm(8),
  prob1_positive = prob1_samp > 0,
  prob1_logic = prob1_positive > 0,
  prob1_char = c("one", "two", "three", "four", "five", "six", "seven", "eight"), 
  prob1_factor = factor(prob1_char, levels = c("low", "moderate", "high")),
)
```

Here's a **code chunk** that is attempting to take the mean of each variable in the data frame above:
```{r}
mean(pull(hw1_df, prob1_samp))
```
What works: The result is a numeric value= 0.05569913
What doesn't: There is no result for non-numeric variables?

Here's a **code chunk** is attempting to convert variable types:
## Logical to numeric
```{r}
coerce_log_num = as.numeric(pull(hw1_df, prob1_logic))
```
## Character to numeric
```{r}
coerce_char_num = as.numeric(pull(hw1_df, prob1_char))
```
## Factor to numeric
```{r}
coerce_fac_num = as.numeric(pull(hw1_df, prob1_factor))
```
* What happens and why: Logical to numeric became binary 1s and 0s/ Character to numeric became all "NA". Factor to numeric also became all "NA".
* Does this explain what happenned to the mean before: The mean was being calculated from the binary coded variables of 1 and 0. 

Here's a second **code chunk** converting and manipulating the variables further:
## converting the logical vector to numeric and multiplying the random sample by the result
```{r}
coerce_log_num*pull(hw1_df, prob1_samp)
```
## converting the logical vector to a factor and multiplying the random sample by the result
```{r}
coerce_log_fac = as.factor(pull(hw1_df, prob1_logic))
  coerce_log_fac*pull(hw1_df, prob1_samp)
```
## converting the logical vector to a factor and then to numeric and multiplying the random sample by the result
```{r}
coerce_log_fac_num = as.numeric(pull(hw1_df, coerce_log_fac)) 
  coerce_log_fac_num*pull(hw1_df, prob1_samp)
```

#__Problem 2:__
Here's a **code chunk** creating a new data frame:
```{r}
prob2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  prob2_logic = x + y > 1,
  log_to_num = as.numeric(prob2_logic),
  log_to_fac = as.factor(prob2_logic)
)
```

PASTE IN-LINE CODE BACK IN HERE:

Here's a **code chunk** of a scatterplot for y vs. x with logical color points:
```{r}
ggplot(prob2_df,aes(x=x, y=y, colors=prob2_logic)) + geom_point()
```



This one has numeric color points:
```{r, eval= false}
ggplot(prob2_df, aes(x = x, y = y)) + geom_point(log_to_num)
```
This one has factorr color points:
```{r, eval= false}
ggplot(prob2_df, aes(x = x, y = y)) + geom_point(log_to_fac)
```

Finally export the first scatterplot to the project directory using ggsave.
