---
title: "Data Science HW1"
author: "Laura Lynch"
date: "9/17/2019"
output: html_document
---

```{r}
library(tidyverse)
```

# __Problem 1:__ 

## Here's a code chunk that samples from a normal distribution:

```{r}
hw1_df = tibble(
  prob1_samp = rnorm(8),
  prob1_positive = prob1_samp > 0,
  prob1_logic = prob1_positive > 0,
  prob1_char = c("one", "two", "three", "four", "five", "six", "seven", "eight"), 
  prob1_factor = factor(prob1_char, levels = c("low", "moderate", "high")),
)
```

## Here's a code chunk that is attempting to take the mean of each variable in the data frame above:

```{r}
mean(pull(hw1_df, prob1_samp))
mean(pull(hw1_df, prob1_positive))
mean(pull(hw1_df, prob1_logic))
mean(pull(hw1_df, prob1_char))
mean(pull(hw1_df, prob1_factor))
```
* prob1_samp WORKS
* prob1_positive
* prob1_logic 
* prob1_char 
* prob1_factor

## Here's a code chunk is attempting to convert variable types:

* Logical to numeric

```{r}
coerce_log_num = as.numeric(pull(hw1_df, prob1_logic))
```

* Character to numeric

```{r, eval= FALSE}
coerce_char_num = as.numeric(pull(hw1_df, prob1_char))
```

* Factor to numeric

```{r}
coerce_fac_num = as.numeric(pull(hw1_df, prob1_factor))
```

What happens and why: Logical to numeric became binary 1s and 0s/ Character to numeric became all "NA". Factor to numeric also became all "NA".
Does this explain what happenned to the mean before: The mean was being calculated from the binary coded variables of 1 and 0. 


## Here's a second code chunk converting and manipulating the variables further:

* converting the logical vector to numeric and multiplying the random sample by the result

```{r}
coerce_log_num*pull(hw1_df, prob1_samp)
```

* converting the logical vector to a factor and multiplying the random sample by the result

```{r}
coerce_log_fac = as.factor(pull(hw1_df, prob1_logic))
  coerce_log_fac*pull(hw1_df, prob1_samp)
```

Result: Not meaningful factors, all are NA

* converting the logical vector to a factor and then to numeric and multiplying the random sample by the result

```{r}
coerce_log_fac = as.factor(pull(hw1_df, prob1_logic))
  coerce_fac_num = as.numeric(coerce_log_fac)
  coerce_fac_num*pull(hw1_df,  prob1_samp)
```

Result: Get numeric variables


# __Problem 2:__

## Here's a code chunk creating a new data frame:

```{r}
prob2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  prob2_logic = x + y > 1,
  log_to_num = as.numeric(prob2_logic),
  log_to_fac = as.factor(prob2_logic)
)
```

Descriptive statistics of dataset "prob2_df":
* There are `r nrow(prob2_df)` rows and `r ncol(prob2_df)` columns in the dataset.
* The mean of x is `r mean(pull(prob2_df, x))`
* The median of the x is `r median(pull(prob2_df, x))`
* The standard deviation of the x is `r median(pull(prob2_df, x))`
* The proportion of cases f for which x + y >1 is `r sum(pull(prob2_df, log_to_num))/(nrow(prob2_df))`

## Scatterplots

* Here's a code chunk of a scatterplot for y vs. x with logical color points:

```{r}
ggplot(prob2_df, aes(x = x, y = y)) + geom_point(aes (color = prob2_logic)) + labs(title ="X vs. Y by Logic Function x + y >1")
ggsave("logic_plot.png", width = 5, height = 5)
```

* This one has numeric color points:

```{r}
ggplot(prob2_df, aes(x = x, y = y)) + geom_point(aes (color = log_to_num)) + labs(title ="X vs. Y by Numeric Variable")
```

* This one has factor color points:

```{r}
ggplot(prob2_df, aes(x = x, y = y)) + geom_point(aes (color = log_to_fac)) + labs(title ="X vs. Y by Factor Variable")
```