---
title: "p8105_hw1_ll3255"
author: "LML"
date: "9/16/2019"
output: html_document
---

```{r}
library(tidyverse)
```

#Problem 1: 
Here's a **code chunk** that samples from a normal distribution:
```{r}
hw1_df = tibble(
  prob1_samp = rnorm(8),
  prob1_positive = prob1_samp > 0,
  prob1_logic = prob1_positive > 0,
  prob1_char = c("one", "two", "three", "four", "five", "six", "seven", "eight"), 
  prob1_factor = factor(prob1_char, levels = c("low", "moderate", "high")),
)
```

Here's a **code chunk** that is attempting to take the mean of each variable in the data frame above:
```{r}
mean(pull(hw1_df, prob1_samp))
```
What works:
What doesn't:

Here's a **code chunk** is attempting to convert variable types:
## Logical to numeric
## Character to numeric
## Factor to numeric
```{r}
coerce_log_num = as.numeric(pull(hw1_df, prob1_logic))
coerce_char_num = as.numeric(pull(hw1_df, prob1_char))
coerce_fac_num = as.numeric(pull(hw1_df, prob1_factor))
```
What happens and why:
Does this explain what happenned to the mean before:

Here's a second **code chunk** converting and manipulating the variables further:
```{r}
coerce_log_num*prob1_samp

coerce_log_fac = as.factor(pull(hw1_df, prob1_logic))
  coerce_log_fac*prob1_samp

coerce_log_fac_num = as.numeric(pull(hw1_df, coerce_log_fac)) 
  coerce_log_fac_num*prob1_samp
```

#Problem 2
Here's a **code chunk** creating a new data frame:
```{r}
prob2_df = tibble(
  x = rnorm(500),
  y = rnorm(500),
  prob2_logic = x + y > 1,
  log_to_num = as.numeric(pull(prob2_df, prob2_logic)),
  log_to_fac = as.factor(pull(prob2_df, prob2_logic))
)
```

Descriptive statistics of dataset `prob2_df`:
* There are `r nrow(prob2_df)` rows and `r ncol(prob2_df)` columns in the dataset.
* The mean of x is `r mean(pulld(prob2_df, x))`
* The median of the x is `r median(pull(prob2_df, x))`
* The standard deviation of the x is `r median(pull(prob2_df, x))`
* The proportion of cases f for which x + y >1 is `r median(pull(prob2_df, prob2_logic))`

Here's a **code chunk** of a scatterplot for y vs. x with logical color points:
```{r}
ggplot(plot_df, aes(x = x, y = y)) + geom_point()
```
This one has numeric color points:
```{r}
ggplot(plot_df, aes(x = x, y = y)) + geom_point()
```
This one has factorr color points:
```{r}
ggplot(plot_df, aes(x = x, y = y)) + geom_point()
```

Finally export the first scatterplot to the project directory using ggsave.
